\section{Experiments}
\label{sec:experiments}
This section is on data analysis by detailing the three main steps described in section \ref{sec:introduction}. The main goal is to find the differences of CST in a healthy brain and a ALS (Amyotrophic Lateral Sclerosis) diseased brain based on quantification some metric values. 

%definition of ALS
%copy and paste from ~\cite{horsfield2002applications}
ALS, also known as motor
neurone disease and Lou Gehrig’s disease, is a fatal,
progressive neuromuscular disease that attacks the motor
neurons controlling voluntary movement. The result is
wasting and atrophy of muscles, leading to difficulties in
speaking or swallowing, stumbling, permanent fatigue
and cramping, amongst other symptoms. The frequency
is about 6 per 100 000. ALS can be broadly classified into
two groups—bulbar onset and limb onset—depending on
how symptoms first manifest themselves. The first
symptoms of bulbar onset ALS include speaking and
swallowing difficulties, resulting from degeneration of
motor neurons concentrated in the brain stem (cortico-
bulbar region). Limb onset patients experience first
symptoms of ALS as muscle weakness in either the arms
or legs. Limb onset patients eventually experience bulbar
symptoms as the disease progresses.
The first study of ALS using diffusion-weighted
imaging was by Segawa et al.26 who, although observing
differences in T2—weighted images in ALS patients (i.e.
hyperintense lesions), did not find any significant
diffusion changes in the posterior limb of the internal
capsule of ALS patients. Wu et al.27 demonstrated
hyperintensity in diffusion-weighted images (signifying
reduced diffusivity) in the cortico-spinal tracts in 11 of 12
ALS patients and five of 12 controls. However, since
diffusion-weighting was only applied in one direction
(left–right) it is likely that this study was confounded by
the effects of diffusion anisotropy.~\cite{horsfield2002applications}


\input{reconstruction}

\input{segmentation}

\input{quantification}


