\begin{abstract}
%why is it relevant - important?
%Why is it critic - difficult?
Nowadays, the large datasets become more and more common. However, traditional visualization techniques, which although allow to visually analyze and explore data, can not scale well with the large one. This restrains the ability of detecting, recognizing and classifying phenomena of interest, such as patterns, clusters, trends, etc.
%What is our goal?
This paper proposes a method for interactive multi-resolution visualization to overcome the problem of traditional visualization techniques when working with a large dataset via hierarchical clustering. Based on hierarchical clustering, users can not only examine the dataset at different levels of detail, but also can explore many regions of interest. The basic idea underlying this method is to choose multiple scales from the hierarchical tree for representing the data at different levels of abstraction, which creates an easy environment for interactive exploration without re-run the clustering algorithm. Moreover, we also define a criterion for evaluating the multiple scale representation based on the concept of \emph{split factor}. An experiment of applying the proposed method into the task of interactive visualization in a clinical case study of the large dMRI (diffusion magnetic resonance imaging) data is also carried out. The results show that our proposed method efficiently provides a friendly tool for visualization the large data.  %\textit{tractography segmentation} in the 
%This paper proposes a method for interactive multi-resolution visualization to overcome the problem of traditional visualization techniques when working with a large dataset by choosing multiple abstraction-levels for representing data via hierarchical clustering. Based on hierarchical clustering, users can not only examine the dataset at different levels of detail, but also can explore many regions of interest. The basic idea underlying this method is to choose multiple scales from hierarchical tree for representing the data at different levels of abstraction, which creates an easy environment for interactive exploration without re-run the clustering algorithm. We evaluate the proposal method by applying it into the task of interactive visualization the large dMRI (diffusion magnetic resonance imaging) data. The results show that our proposed method efficiently provides a friendly tool for visualization the large data.  %\textit{tractography segmentation} in the 
\end{abstract}

